Perturbation Theory in the Complex Plane: Exceptional Points and Where to Find Them

### 2020-11-30 Manuscript/EPAWTFT.tex
Added discussion on UMP relationship to critical point

\newcommand{\Lup}{\mathcal{L}^{\uparrow}}
\newcommand{\Ldown}{\mathcal{L}^{\downarrow}}
\newcommand{\Lsi}{\mathcal{L}^{\sigma}}
\newcommand{\Rup}{\mathcal{R}^{\uparrow}}
\newcommand{\Rdown}{\mathcal{R}^{\downarrow}}
\newcommand{\Rsi}{\mathcal{R}^{\sigma}}

@@

%==========================================%
\subsection{Spin-Contamination in the Hubbard Dimer}
%==========================================%

@@

>>>>

% Making life easier
\newcommand{\Lup}{\mathcal{L}^{\uparrow}}
\newcommand{\Ldown}{\mathcal{L}^{\downarrow}}
\newcommand{\Lsi}{\mathcal{L}^{\sigma}}
\newcommand{\Rup}{\mathcal{R}^{\uparrow}}
\newcommand{\Rdown}{\mathcal{R}^{\downarrow}}
\newcommand{\Rsi}{\mathcal{R}^{\sigma}}
\newcommand{\vhf}{v_{\text{HF}}}

@@

%==========================================%
\subsection{Spin-Contamination in the Hubbard Dimer}
\label{sec:spin_cont}
%==========================================%

@@

%------------------------------------------------------------------%
% Figure on the UMP critical point
%------------------------------------------------------------------%
\begin{figure*}[t]
\begin{subfigure}{0.49\textwidth}
\includegraphics[height=0.65\textwidth]{ump_critical_point}
\subcaption{\label{subfig:ump_cp}}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
\includegraphics[height=0.65\textwidth]{ump_ep_to_cp}
\subcaption{\label{subfig:ump_ep_to_cp}}
\end{subfigure}
\caption{%
\hugh{%
    The UMP ground-state EP becomes a critical point in the strong correlation limit (large $U/t$).
    (\subref{subfig:ump_cp}) As the $U/t$ increases, the avoided crossing on the real $\lambda$ axis
    becomes increasingly sharp.
    (\subref{subfig:ump_ep_to_cp}) The convergence of the EPs onto the real axis for $U/t \rightarrow \infty$
    mirrors the formation of the RMP critical point and other QPTs in the infinite basis set limit.
}
\label{fig:UMP_cp}}

\end{figure*}
%------------------------------------------------------------------%

% RELATIONSHIP BETWEEN QPT AND UMP
\hugh{%
In Section~\ref{sec:spin_cont} we showed that the slow convergence of the UMP series for the strongly correlated
Hubbard dimer was due to a complex-conjguate pair of  EPs just outside the radius of convergence.
These EPs have positive real components and small imaginary components (Fig.~\ref{fig:UMP}), suggesting a potential 
connection to MP critical points and QPTs.
For $\lambda>1$, the HF potential becomes an attractive component in Stillinger's 
Hamiltonian [Eq.~\eqref{eq:HamiltonianStillinger}], while the explicit electron-electron interaction
becomes increasingly repulsive.
Critical points along the positive real $\lambda$ axis for closed-shell molecules then represent 
points were the two-electron repulsion overcomes the attractive HF potential and an electron
are successively expelled from the molecule.\cite{Sergeev_2006}
}

\hugh{%
Symmetry-breaking in the UMP reference creates different HF potentials for the spin-up and spin-down electrons.
Consider the reference UHF solution where the spin-up and spin-down electrons are localised on the left and
right sites respectively.  
The spin-up HF potential will then be a repulsive interaction from the spin-down electron 
density that is centred around the right site (and vice-versa).
As $\lambda$ becomes greater than 1 and the HF potentials become attractive, there will be a sudden
driving force for the electrons to swap sites.
This swapping process can also be represented as a double excitation, and thus an avoided crossing will occur
for $\lambda \geq 1$ (Fig.~\ref{subfig:ump_cp}).
Note that, although this appears to be an avoided crossing between the ground and first-excited state, 
the earlier excited-state avoided crossing means that the first-excited state qualitatively 
represents the double excitation for $\lambda > 0.5.$
}

% SHARPNESS AND QPT
\hugh{%
The ``sharpness'' of the avoided crossing is controlled by the correlation strength $U/t$.
For small $U/t$, the HF potentials will be weak and the electrons will delocalise over the two sites,
both in the UHF reference and as $\lambda$ increases.
This delocalisataion dampens the electron swapping process and leads to a ``shallow'' avoided crossing
that corresponds to EPs with non-zero imaginary components (solid lines in Fig.~\ref{subfig:ump_cp}).
As $U/t$ becomes larger, the HF potentials become stronger and the on-site repulsion dominates the hopping
term to make electron delocalisation less favourable.
These effects create a stronger driving force for the electrons to swap sites until eventually this swapping
occurs exactly at $\lambda = 1$.
In this limit, the ground-state EPs approach the real axis (Fig.~\ref{subfig:ump_ep_to_cp}) and the avoided 
crossing creates a gradient discontinuity in the ground-state energy (dashed lines in Fig.~\ref{subfig:ump_cp}).
We therefore find that, in the strong correlation limit, the symmetry-broken ground-state EP becomes becomes
a new type of MP critical and represents a QPT in the perturbation approximation. 
Furthermore, this argument explains why the dominant UMP singularity lies so close, but always outside, the 
radius of convergence (see Fig.~\ref{fig:RadConv}).
}
